{\bf Practical Concurrency Testing} \\ \normalsize \vspace{1em} \begin{tabular}{c} {\em or: How I Learned to Stop Worrying and Love the Exponential Explosion} \end{tabular}

### 2018-12-19 thesis/landslide.tex
quicksand impl
proofread text wrapping, etc in landslide

						(\cref{sec:landslide-foundabug}, \cref{sec:landslide-bugreport})

@@

					\item Select which thread should run next (\cref{sec:landslide-arbiter}),

@@

						(\cref{sec:landslide-interrupce},
						\cref{sec:landslide-scheduler-interrupce}).

@@

				(\cref{sec:landslide-interrupce}, \cref{sec:landslide-scheduler-interrupce}),

@@

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\cref{sec:landslide-icb})
	\item {\tt -0}: enable Preempt-Everywhere mode (\cref{sec:quicksand-eval}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\cref{sec:tm-eval}, incompatible with {\tt -C}).
		According to \cref{sec:quicksand-soundness}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (\Cref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\cref{sec:tm-implementation});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\cref{sec:tm-implementation})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\cref{sec:tm-implementation})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
\end{itemize}

@@

\begin{itemize}
	\item {\tt mem\_check\_shared\_access()} ({\tt memory.c})
		will assert that kernel memory is direct-mapped.
	\item {\tt use\_after\_free()} ({\tt memory.c})
		will ignore use-after-free reads originating from kernel code during the {\tt swexn} system call
		(an extremely common and neither harmful nor technically interesting bug among student implementations).
	\item {\tt cause\_test()} ({\tt test.c})
		will issue keyboard input to type the test case name and press enter
		when the initialization sequence completes and the shell is blocked on {\tt readline}.
	\item {\tt kern\_mutex\_block\_others()} ({\tt schedule.c})
		will handle the special ``blocked on via mutex'' state changes
		whenever a mutex is acquired or released,
		for kernels which use the {\tt tell\_landslide\_kern\_mutex\_blocking()} annotation.
	\item {\tt sched\_update\_kern\_state\_machine()} ({\tt schedule.c})
		will handle the reference kernel's invocation of {\tt sched\_unblock()} within {\tt cond\_\allowbreak{}signal()}
		as a signal event for happens-before analysis.
	\item {\tt cause\_timer\_interrupt\_immediately()} ({\tt x86.c})
		will read the {\tt esp0} value out of the TSS to support user-to-kernel mode switches
		in timer interrupts injected during userspace execution.
	\item {\tt splice\_pre\_vanish\_trace} ({\tt stack.c})
		will, when a {\tt vanish}ing thread has already
		\revisionminor{unmapped all of its user memory,}
		%deleted its Simics symbol table object,
		splice in a saved ``pre-vanish'' stack trace
		(saved previously in {\tt sched\_update\_kern\_state\_machine()})
		so the user can see the userspace execution sequence preceding the {\tt vanish} invocation.
	\item {\tt stack\_trace} ({\tt stack.c})
		will, when {\tt ebp} crosses from kernel- to userspace across a system call boundary
		(a reference kernel feature to allow Simics stack traces to cross same),
		use {\tt PATHOS\_SYSCALL\_IRET\_DISTANCE} (\cref{sec:landslide-config-landslide})
		to track {\tt esp}'s value across the stack switch.
\end{itemize}

@@

\begin{itemize}
	\item {\tt suspected\_data\_race()}
	\item User or kernel {\tt mutex\_lock()} or {\tt mutex\_unlock()} call
	\item Custom kernel preemption point using {\tt tell\_landslide\_preempt()}
		(relic of \cite{landslide}, largely obsoleted by data-race preemption pooints)
\end{itemize}

@@

Landslide detects such ad-hoc {\em yield-loop blocking}
by keeping a counter for each thread to track how many {\tt yield}s it has invoked since the last ``interesting'' activity.
% nb. not system calls, not even d/mr, as some student mutexes may use those and you don't want to count them
``Interesting'' here is defined heuristically as any other known P2 API call,
with the exception of {\tt mutex\_lock()} and {\tt mutex\_unlock()}.%
\footnote{Landslide must recognize yield-blocking loops that contain mutex operations
to allow for open-coded reimplementations of {\tt cond\_wait()},
% see comment above "USER_MUTEX_YIELD_ACTIVITY" in user sync.c
which for example the {\tt paraguay} test uses intentionally,
because it is testing the correctness of the student's {\tt cond\_wait()}.}
Whenever this counter reaches the heuristic cutoff of 10 ({\tt TOO\_MANY\_YIELDS}),
Landslide declares the thread blocked,
and treats it just as if it had invoked {\tt deschedule()} for purposes of DPOR
({\tt check\_user\_yield\_activity()} in {\tt user\_sync.c}).
% this also involves propagating yield-blockedness back up the branch, along those 10 pps where it was still counting up to 10
% and also refusing to run any other threads while yield is going on - KEEP_RUNNING_YIELDING_THREADS
%
In cases where {\tt yield} itself is not involved,
Landslide also counts the number of atomic instructions ({\tt xchg}, {\tt xadd}, {\tt cmpxchg}, et cetera),
and marks the thread blocked if it exceeds 100 such ({\tt TOO\_MANY\_XCHGS\_TIGHT\_LOOP}).%
\footnote{If preemption points exist in between, instead only 20 such ({\tt TOO\_MANY\_XCHGS\_WITH\_PPS}),
to avoid stressing DPOR's $O(n)$ independence computation.}
When such a loop contains neither {\tt yield} nor atomics,
Landslide falls back on the standard infinite loop detector and reports a bug directly, as described above.
Future work could extend this to include more modern ways of establishing memory safety such as
acquire/release barriers
%\cite{sully-thesis}
and hardware transactions.

>>>>

						(\cref{sec:landslide-bugreport},
						\cref{sec:landslide-foundabug})

@@

					\item Select which thread should run next (\cref{sec:landslide-arbiter}).

@@

						(\cref{sec:landslide-scheduler-interrupce},
						\cref{sec:landslide-interrupce}).

@@

				(\cref{sec:landslide-scheduler-interrupce},
				\cref{sec:landslide-interrupce}),

@@

\begin{itemize}
	\item {\tt -C}: enable ``control experiment'' mode, i.e., run only 1 instance of Landslide, with all (non-data-race) preemption points enabled in advance
	\item {\tt -I}: enable Iterative Context Bounding (requires {\tt -C}, although future work may relax this restriction);
		this generally causes bugs to be found faster should they exist, but degrades completion time
		(\cref{sec:landslide-icb})
	\item {\tt -0}: enable Preempt-Everywhere mode
		(\cref{sec:quicksand-impl-modes}, \cref{sec:quicksand-impl-landslide}, requires {\tt -C})
	\item {\tt -M}: enable Maximal State Space mode,
		which prioritizes the maximal state space to optimize for fast verification,
		abandoning all subset jobs even if they might find bugs faster
		(\cref{sec:tm-eval}, incompatible with {\tt -C}).
		According to \cref{sec:quicksand-soundness}'s soundness proofs,
		this is equivalent to {\tt -0}
		(and according to my experience, \revisionminor{much} faster as well).
	\item {\tt -H}: use Limited Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for Pebbles kernelspace mode)
	\item {\tt -V}: use vector-clock-based Pure Happens-Before for data-race analysis (\cref{sec:background-hb})
		(default for P2/PSU userspace and Pintos modes)
	\item {\tt -X}: support transactional memory (\Cref{chap:tm})
	\item {\tt -A}: support multiple abort codes during transaction failure (\cref{sec:tm-implementation});
		required for testing programs which behave differently under different abort circumstances,
		but impacts the state space size
	\item {\tt -S}: suppress retry aborts during transaction failure (\cref{sec:tm-implementation})
	\item {\tt -R}: enable retry-set state space reduction for transactional tests (\cref{sec:tm-implementation})
	\item {\tt -P}: support Pintos architecture (enabled automatically when {\tt pintos-setup.sh} is run)
	\item {\tt -4}: support Pebbles architecture (enabled automatically when either {\tt p2-setup.sh} or {\tt psu-setup.sh} is run)
	\item {\tt -e ETAFACTOR}: configure heuristic state space ETA deferring factor (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
	\item {\tt -E ETATHRESH}: configure heuristic threshold of state space progress for judging ETA stability (\cref{sec:quicksand-id}; described in detail in {\tt id/option.c})
\end{itemize}

@@

\begin{itemize}
	\item {\tt mem\_check\_shared\_access()} ({\tt memory.c})
		will assert that kernel memory is direct-mapped.
	\item {\tt use\_after\_free()} ({\tt memory.c})
		will ignore use-after-free reads originating from kernel code during the {\tt swexn} system call
		(an extremely common and neither harmful nor technically interesting bug among student implementations).
	\item {\tt cause\_test()} ({\tt test.c})
		will issue keyboard input to type the test case name and press enter
		when the initialization sequence completes and the shell is blocked on {\tt readline}.
	\item {\tt kern\_mutex\_block\_others()} ({\tt schedule.c})
		will handle the special ``blocked on via mutex'' state changes
		whenever a mutex is acquired or released,
		for kernels which use the {\tt tell\_landslide\_kern\_mutex\_blocking()} annotation.
	\item {\tt sched\_update\_kern\_state\_machine()} ({\tt schedule.c})
		will handle the reference kernel's invocation of {\tt sched\_unblock()} within {\tt cond\_\allowbreak{}signal()}
		as a signal event for happens-before analysis.
	\item {\tt cause\_timer\_interrupt\_immediately()} ({\tt x86.c})
		will read the {\tt esp0} value out of the TSS to support user-to-kernel mode switches
		in timer interrupts injected during userspace execution.
	\item {\tt splice\_pre\_vanish\_trace()} ({\tt stack.c})
		will, when a {\tt vanish}ing thread has already
		\revisionminor{unmapped all of its user memory,}
		%deleted its Simics symbol table object,
		splice in a saved ``pre-vanish'' stack trace
		(saved previously in {\tt sched\_update\_kern\_state\_machine()})
		so that the user can see the userspace execution sequence preceding the {\tt vanish} invocation.
	\item {\tt stack\_trace()} ({\tt stack.c})
		will, when {\tt ebp} crosses from kernel- to userspace across a system call boundary
		(a reference kernel feature to allow Simics stack traces to cross same),
		use {\tt PATHOS\_SYSCALL\_IRET\_DISTANCE} (\cref{sec:landslide-config-landslide})
		to track {\tt esp}'s value across the stack switch.
\end{itemize}

@@

\begin{itemize}
	\item {\tt suspected\_data\_race()}
	\item User or kernel {\tt mutex\_lock()} or {\tt mutex\_unlock()} call
	\item Custom preemption point requested by the user with {\tt tell\_landslide\_preempt()}
		(relic of \cite{landslide}, largely obsoleted by data-race preemption points (\cref{sec:quicksand-dr-pps}))
\end{itemize}

@@

Landslide detects such ad-hoc {\em yield-loop blocking}
by keeping a counter for each thread to track how many {\tt yield}s it has invoked since the last interesting activity.
% nb. not system calls, not even d/mr, as some student mutexes may use those and you don't want to count them
``Interesting'' here is defined heuristically as any other known P2 API call,
with the exception of {\tt mutex\_lock()} and {\tt mutex\_unlock()}.%
\footnote{Landslide must recognize yield-blocking loops that contain mutex operations
to allow for open-coded reimplementations of {\tt cond\_wait()},
% see comment above "USER_MUTEX_YIELD_ACTIVITY" in user sync.c
which for example the {\tt paraguay} test uses intentionally,
because it is testing the correctness of the student's {\tt cond\_wait()}.}
Whenever this counter reaches the heuristic cutoff of 10 ({\tt TOO\_MANY\_YIELDS}),
Landslide declares the thread blocked,
and treats it just as if it had invoked {\tt deschedule()} for purposes of DPOR
({\tt check\_user\_yield\_activity()} in {\tt user\_sync.c}).
% this also involves propagating yield-blockedness back up the branch, along those 10 pps where it was still counting up to 10
% and also refusing to run any other threads while yield is going on - KEEP_RUNNING_YIELDING_THREADS
%
In cases where {\tt yield} itself is not involved,
Landslide also counts the number of atomic instructions ({\tt xchg}, {\tt xadd}, {\tt cmpxchg}, et cetera),
and marks the thread blocked if it exceeds 100 such ({\tt TOO\_MANY\_XCHGS\_TIGHT\_LOOP}).%
\footnote{If preemption points exist in between, instead only 20 such ({\tt TOO\_MANY\_XCHGS\_WITH\_PPS}),
to avoid stressing DPOR's $O(n)$ independence computation.}
When such a loop contains neither {\tt yield} nor atomics,
Landslide falls back on the standard infinite loop detector and reports a bug directly, as described above.
Future work could extend this to include more modern ways of establishing memory safety such as
acquire/release barriers
%\cite{sully-thesis}
and hardware transactions.
